### 2020-05-06 main.Rmd
corrected minor mistake in spelling and added the path of non-effector data

The performance of the trained classifiers is dependent on the quality of the input training data, so it was important that as high quality set of annotated effectors as possible was collected. To this end we used PHI-Base [@phibase] as our primary sequence store. Sequences in PHI-Base are human curated from the literature and have therefore been noted in experimental studies. They do not derive from large scale annotations or contain hypothetical or predicted proteins. This attribute makes it ideal for our purposes as the effectors in PHI-Base are those that have been specifially reported as such in the published literature and are not of the class of sequences that are merely suspected of being effectors on the basis of carrying a secretion signal. To collect effector sequences we parsed a whole database text dump of version 4.8 [https://github.com/PHI-base/data](https://github.com/PHI-base/data/blob/master/releases/phi-base_current.csv) ^[Ruth, who provided the text dump, was it Martin Urban? Where is the file in your repos?], all proteins marked as plant pathogen effectors were filtered with IDs and UniProt IDs. These ID then were mapped to PHI-Base fasta. If the sequences are not available in PHI-Base fasta data, then the sequence retrieved from UniProt. The sequences retrieved can be seen in the data file in this manuscript's repository [(`effector_data.csv`)](https://github.com/TeamMacLean/ruth-effectors-prediction/blob/master/data/getting-data-new/binary-class-data/effector_data.csv). Effector sequences were then divided into taxonomic groups as bacterial, fungal or oomycete derived accordingly. The total number of plant effectors per group can be seen in Table \@ref(tab:effsummary). The species and effector count in each group can be seen in Tables \@ref(tab:bacefftable), \@ref(tab:fungiefftable) and \@ref(tab:oomefftable)

Sequences for non-effector, secreted proteins were collected using a similar pipeline. Randomly selected proteins from each species carrying secretion signals were extracted from Ensembl databases using the BioMart tool. For each species noted in Tables \@ref(tab:bacefftable), \@ref(tab:fungiefftable) and \@ref(tab:oomefftable) we collected an identical number of non-effector, secreted proteins to that collected in the effector set. This gave us a balanced data set of effector proteins as positive learning examples and non-effector secreted proteins as negative learning examples. Figure \@ref(fig:noneffectorcollection) summaries the process of getting non-effector set. Using the gene information obtained from effector data, we downloaded secreted sequences from the same genome from each individual effector sequence depending on the availability of them in Ensembl Fungi and Ensembl Protists, otherwise we downloaded sequences from the same species as the individual effector using BioMart tool. However, BioMart tool is not supported for Ensembl Bacteria. Therefore, we downloaded the bacteria sequences using Ensmbl Bacteria FTP. Additionaly, we only selected the secreted sequences from prediction results of SignalP ran on the downloaded data. We then blast the combination of both secreted sequences data downloaded manually using BioMart tool and using Ensembl FTP with the effector sequence data, and removed identical sequences. This process is done in order to make sure there is no effector data in our candidate non-effector sample. Last step in this process then was to sample an identical number of non-effector data as in Tables \@ref(tab:bacefftable), \@ref(tab:fungiefftable) and \@ref(tab:oomefftable). In this step, we also blast every time we took a sample, so that there is no identical sequences within non-effector set.

>>>>

The performance of the trained classifiers is dependent on the quality of the input training data, so it was important that as high quality set of annotated effectors as possible was collected. To this end we used PHI-Base [@phibase] as our primary sequence store. Sequences in PHI-Base are human curated from the literature and have therefore been noted in experimental studies. They do not derive from large scale annotations or contain hypothetical or predicted proteins. This attribute makes it ideal for our purposes as the effectors in PHI-Base are those that have been specifially reported as such in the published literature and are not of the class of sequences that are merely suspected of being effectors on the basis of carrying a secretion signal. To collect effector sequences we parsed a whole database text dump of version 4.8 [https://github.com/PHI-base/data](https://github.com/PHI-base/data/blob/master/releases/phi-base_current.csv), all proteins marked as plant pathogen effectors were filtered with IDs and UniProt IDs. These ID then were mapped to PHI-Base fasta. If the sequences are not available in PHI-Base fasta data, then the sequence retrieved from UniProt. The sequences retrieved can be seen in the data file in this manuscript's repository [(`effector_data.csv`)](https://github.com/TeamMacLean/ruth-effectors-prediction/blob/master/data/getting-data-new/binary-class-data/effector_data.csv). Effector sequences were then divided into taxonomic groups as bacterial, fungal or oomycete derived accordingly. The total number of plant effectors per group can be seen in Table \@ref(tab:effsummary). The species and effector count in each group can be seen in Tables \@ref(tab:bacefftable), \@ref(tab:fungiefftable) and \@ref(tab:oomefftable)

Sequences for non-effector, secreted proteins were collected using a similar pipeline. Randomly selected proteins from each species carrying secretion signals were extracted from Ensembl databases using the BioMart tool. For each species noted in Tables \@ref(tab:bacefftable), \@ref(tab:fungiefftable) and \@ref(tab:oomefftable) we collected an identical number of non-effector, secreted proteins to that collected in the effector set. This gave us a balanced data set of effector proteins as positive learning examples and non-effector secreted proteins as negative learning examples. Figure \@ref(fig:noneffectorcollection) summaries the process of getting non-effector set. Using the gene information obtained from effector data, we downloaded secreted sequences from the same genome from each individual effector sequence depending on their availability on Ensembl Fungi and Ensembl Protists, otherwise we downloaded sequences from the same species as the individual effector. BioMart tool is used in this process.  However, BioMart tool is not supported for Ensembl Bacteria. Therefore, we downloaded the bacteria sequences using Ensembl Bacteria FTP. Additionaly, we only selected the secreted sequences from prediction results of SignalP ran on the downloaded data. We then blast the combination of both secreted sequences data downloaded manually using BioMart tool and using Ensembl FTP with the effector sequence data, and removed identical sequences. This process is done in order to make sure there is no effector data in our candidate non-effector sample. Last step in this process then was to sample an identical number of non-effector data as in Tables \@ref(tab:bacefftable), \@ref(tab:fungiefftable) and \@ref(tab:oomefftable). In this step, we also ran blast every time we took a sample, so that there is no identical sequences within non-effector set. The non-effector set sequences obtained can be seen in the data file in this manuscript's repository [(`non_effector_data.csv`)](https://github.com/TeamMacLean/ruth-effectors-prediction/tree/master/data/secreted_data).
